---
title: "An introduction to the package RepoTime"
author: "David Salgado"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## A notation for usual time intervals

In statistical production it is common to refer to natural time intervals such as
months, terms, semesters, years, etc. We will use a compact notation to denote
them according to the following convention:

**Interval** | **Notation**
-----------|--------------------------
Fortnight  | Q (*Quincena* in Spanish)
Month      | M
Bimonth    | B
Term       | T
Semester   | S
Year       | A (*Año* in Spanish)

Occasionally in statistical operations it is necessary to rotate the sample. 
This drives us to identify those data sets with a rotated sample from those 
without a rotation. We modify the preceding table by introducing a second letter
with self-evident meaning:

**Interval** | **Notation**
-------------|-------------------------------
Fortnight    | QQ, QR (*Quincena* in Spanish)
Month        | MM, MR
Bimonth      | BB, BR
Term         | TT, TR
Semester     | SS, SR
Year         | AA, AR (*Año* in Spanish)

Now each time interval is denoted with the following notation:

**Interval** | **Initials** | **Notation**
-------------|--------------|-----------------------------------------------------------
Fortnight    | QQ, QR       | *qmmyyyy*, *q*=1,2; *mm*= 01, ..., 12; *yyyy*= 4-digit year 
Month        | MM, MR       | *mmyyyy*, *mm* = 01, ..., 12; *yyyy* = 4-digit year
Bimonth      | BB, BR       | *byyyy*, *b* = 1, ..., 6; *yyyy* = 4-digit year
Term         | TT, TR       | *tyyyy*, *t* = 1, ..., 4; *yyyy* = 4-digit year
Semester     | SS, SR       | *syyyy*, *s* = 1, 2; *yyyy* = 4-digit year
Year         | AA, AR       | *yyyy* = 4-digit year

A few examples:

* QQ1032015: first half of March, 2015.
* MM101999: October, 1999.
* BR22000: March and April, 2000 with a rotated sample with respect to January and February, 2000.
* TR32014: Third term (July, August, September), 2014.
* SR11992: First semester (Januey to June), 1992 with a rotated sample with respect to the second semester of 1991.
* AA2002: Year 2002.

## The class RepoTimeInt

To implement this notation we have defined a class named `RepoTimeInt` with two attributes:

* `Repo`: a character vector with intervals expressed in the above notation.
* `lubriInt`: a list of object of class `Interval` from the package lubridate.

The attribute `Repo` contains the specification of several time intervals in the former notation whereas the attribute `lubriInt` expresses this same set of time intervals using the class `Interval` from the package lubridate.

The general idea of the package is to express time intervals with the agreed notation but to work internally with the functions already defined for `Interval`-class methods. 

To build an object of class `RepoTimeInt` we have defined a constructor:

```{r}
library(RepoTime)
Months2015 <- newRepoTime(paste0('MM', c(paste0('0', 1:9), 10:12), '2015'))
```

Internally this constructor makes use of the function `RepoTimeTolubri`:

```{r}
RepoTimeTolubri('MM012015')
```

This function has an inverse function:

```{r}
lapply(RepoTimeTolubri(c('MM012015', 'AA2016')), lubriToRepoTime)
```

Los atributos pueden recuperarse con los respectivos *getters*:

```{r}
getRepo(Months2015)
getlubriInt(Months2015)
```

## Operations with time intervals

We have adapted some methods to manipulate `RepoTimeInt`-class objects. Firstly, given an object `RepoTimeInt` with length greater than 1, we can obtain a new interval resulting from the union of all intervals contained in the object:

```{r}
Length(Months2015)
Union(Months2015)
```

If we have two objects of length 1, we can obtain a new period as the union of both time intervals:

```{r}
Union(Months2015['MM012015'], Months2015['MM032015'])
```

Similarly, we can obtain the common time intervals of two given objects:

```{r}
Intersect(newRepoTime('AA2015'), newRepoTime('MM012015'))
```

Or of a set of intervals:

```{r}
Intersect(Months2015)
```

We can also create a sequence of time intervals as a list of objects of class `RepoTimeInt`:

```{r}
Seq(newRepoTime('MM112015'), newRepoTime('MM032016'))
```

This method allows us also to take into account whether to include or not data sets with rotated samples and/or to indicate which time interval contains the rotated sample:

```{r}
Seq(newRepoTime('MM112015'), newRepoTime('MM032016'), Rot = FALSE)
Seq(newRepoTime('TT32015'), newRepoTime('TT12016'), Rot = TRUE, RotPer = '4')
```

Notice that to create a sequence of rotated time intervals, we must do it interval by interval:

```{r}
newRepoTime(paste0('TR', 1:4, '2015'))
```

Finally, we can add duration objects to a time interval:

```{r}
library(lubridate)
RepoPeriod <- newRepoTime('MM022015')
RepoPeriod + months(1)
RepoPeriod + years(1)
RepoPeriod + weeks(1)
```
